Silenced in the Chamber: Analyzing Gendered Speech Interruptions in the Australian Parliament (1998–2022)

### 2025-03-28 paper/paper.qmd
abstract

---
title: "Silenced in the Chamber: Analyzing Gendered Speech Interruptions in the Australian Parliament (1998–2022)"
author: 
  - Hannah Yu
  - Rohan Alexander
thanks: "Code and data are available at: https://github.com/hannahyu07/Hansard_Interruptions"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The model was implemented in R using the `rstanarm` package [@rstanarm], which applies Markov chain Monte Carlo (MCMC) sampling to estimate the posterior distributions of the model parameters. To optimize computational efficiency, I randomly sampled 1000 observations for model fitting, ensuring a balance between computational feasibility and statistical reliability.

@@

A relationship is expected between gender, party affiliation, and speech interruptions in parliamentary debates. Male speakers are likely to face more interruptions than female speakers, reflecting common patterns in political discourse. However, interruptions may still occur for female speakers depending on the context of their speech.

For party affiliation, the ALP and LNP, as the two main parties, are expected to show different interruption patterns. ALP members may experience more interruptions, especially when in opposition, while LNP members may face fewer interruptions, particularly when in government. Other parties and independents may not follow a clear pattern due to their varying roles in debates.

@@

@tbl-modelresults1 identifies key predictors including gender, party, and word count of speech interruptions in parliamentary proceedings. The intercept, which represents the baseline category in the model, corresponds to female speakers from the LP party. This suggests that all other coefficients should be interpreted as deviations from this baseline.Gender plays an important role, with male speakers more likely to experience interruptions compared to female speakers. This suggests that men face slightly higher odds of interjections during their speeches. Party affiliation also significantly influences interruptions, with some parties, such as the ALP, experiencing more frequent interjections, while others, like the LNP, are interrupted less often. These patterns highlight how political dynamics can shape interactions in debates.

@@

The Bayesian logistic regression model quantifies gender-based interruptions in the Australian Parliament, assessing whether female speakers face disproportionately more interjections than their male counterparts. However, interruptions alone do not fully capture discourse dynamics. To further explore these patterns, I applied Structural Topic Modeling (STM) to identify the most discussed topics in parliamentary speeches, incorporating gender and party affiliation as prevalence covariates to examine how different groups engage with various topics.

By linking the most discussed topics in a given year to the most interrupted gender, I aim to determine whether interruptions are associated with particular types of discourse. For example, if female parliamentarians experienced the highest interruption rates in 2020, I will compare this with the dominant topics that year to uncover potential gendered patterns. This dual approach—quantifying interruptions through regression and analyzing speech content through topic modeling—offers deeper insight into both who gets interrupted and what is being discussed when those interruptions occur.

The process began with data preprocessing, where I filtered the dataset to include only relevant entries, removed missing values, and structured the text data for analysis. A document-feature matrix (DFM) was then created, representing word frequencies while eliminating punctuation, stopwords, and low-frequency terms to improve model efficiency. The STM model was trained with 10 topics and incorporated party affiliation and speaker gender as prevalence covariates, allowing for an examination of how these factors influence topic distribution. The model was optimized over 50 iterations, ensuring convergence and stability. The final output provided a breakdown of the dominant topics, with key terms extracted using highest probability words, FREX words (distinctive to each topic), Lift words (rare but highly relevant), and Score words (best summarizing a topic’s meaning).

@@

To analyze the evolution of parliamentary discourse over time, I ran the Structural Topic Model (STM) on each year from 1998 to 2022, extracting the most prominent topics discussed in Australian parliamentary speeches annually. This approach allows us to track changes in political priorities and public policy discussions across different governments and time periods.

>>>>

---
title: "Silenced in the Chamber: Analyzing Gendered Speech Interruptions in the Australian Parliament (1998–2022)"
author: 
  - Hannah Yu
  - Rohan Alexander
thanks: "Code and data are available at: https://github.com/hannahyu07/Hansard_Interruptions"
date: today
date-format: long
abstract: "Interruptions shape parliamentary discourse, influencing power dynamics and legislative interactions. Using Hansard records from 1998 to 2022, this study applies Bayesian logistic regression and Structural Topic Modeling (STM) to examine how gender, party affiliation, and debate topics affect speech interruptions in the Australian Parliament. The analysis finds that while men experience more interruptions overall, differences between genders diminish when accounting for speech frequency, with political party and debate topics playing a larger role in determining interruption rates. These findings highlight the complex nature of parliamentary speech, emphasizing that interruptions are not solely gender-driven but are also shaped by institutional and political contexts."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The model was implemented in R using the `rstanarm` package [@rstanarm], which applies Markov chain Monte Carlo (MCMC) sampling to estimate the posterior distributions of the model parameters. To optimize computational efficiency, we randomly sampled 1000 observations for model fitting, ensuring a balance between computational feasibility and statistical reliability.

@@

Male speakers are likely to face more interruptions than female speakers, reflecting common patterns in political discourse. However, interruptions may still occur for female speakers depending on the context of their speech.For party affiliation, the ALP and LNP, as the two main parties, are expected to show different interruption patterns. ALP members may experience more interruptions, especially when in opposition, while LNP members may face fewer interruptions, particularly when in government. Other parties and independents may not follow a clear pattern due to their varying roles in debates.

@@

@tbl-modelresults1 identifies key predictors including gender, party, and word count of speech interruptions in parliamentary proceedings. The intercept, which represents the baseline category in the model, corresponds to female speakers from the LP party. This suggests that all other coefficients should be interpreted as deviations from this baseline. Gender plays an important role, with male speakers more likely to experience interruptions compared to female speakers. This suggests that men face slightly higher odds of interjections during their speeches. Party affiliation also significantly influences interruptions, with some parties, such as the ALP, experiencing more frequent interjections, while others, like the LNP, are interrupted less often. These patterns highlight how political dynamics can shape interactions in debates.

@@

The Bayesian logistic regression model quantifies gender-based interruptions in the Australian Parliament, assessing whether female speakers face disproportionately more interjections than their male counterparts. However, interruptions alone do not fully capture discourse dynamics. To further explore these patterns, we applied Structural Topic Modeling (STM) to identify the most discussed topics in parliamentary speeches, incorporating gender and party affiliation as prevalence covariates to examine how different groups engage with various topics.

By linking the most discussed topics in a given year to the most interrupted gender, we aim to determine whether interruptions are associated with particular types of discourse. For example, if female parliamentarians experienced the highest interruption rates in 2020, we will compare this with the dominant topics that year to uncover potential gendered patterns. This dual approach—quantifying interruptions through regression and analyzing speech content through topic modeling—offers deeper insight into both who gets interrupted and what is being discussed when those interruptions occur.

The process began with data preprocessing, where we filtered the dataset to include only relevant entries, removed missing values, and structured the text data for analysis. A document-feature matrix (DFM) was then created, representing word frequencies while eliminating punctuation, stopwords, and low-frequency terms to improve model efficiency. The STM model was trained with 10 topics and incorporated party affiliation and speaker gender as prevalence covariates, allowing for an examination of how these factors influence topic distribution. The model was optimized over 50 iterations, ensuring convergence and stability. The final output provided a breakdown of the dominant topics, with key terms extracted using highest probability words, FREX words (distinctive to each topic), Lift words (rare but highly relevant), and Score words (best summarizing a topic’s meaning).

@@

To analyze the evolution of parliamentary discourse over time, we ran the Structural Topic Model (STM) on each year from 1998 to 2022, extracting the most prominent topics discussed in Australian parliamentary speeches annually. This approach allows us to track changes in political priorities and public policy discussions across different governments and time periods.
